LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment

### 2024-11-14 article/main.tex
capitalize Figure everywhere

\begin{table*}[ht!]
    \setlength\tabcolsep{6pt}
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{@{}llllll@{}}
            \toprule
            LFTK ID         & Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark\\ \toprule
            t\_word         & total words                       & wordsent         & surface          & very concrete   & lexical                 \\
            t\_sent         & total sentences                   & wordsent         & surface          & very concrete   & syntactic               \\ \bottomrule
            n\_uverb        & total number of unique verbs      & partofspeech     & syntax           & concrete        & syntactic               \\
            n\_uadj         & total number of unique adjectives & partofspeech     & syntax           & concrete        & syntactic               \\ \bottomrule
            simp\_ttr       & simple type token ratio           & typetokenratio   & lexico-semantics & regular         & pragmatic               \\
            a\_verb\_pw     & number of verbs per word          & partofspeech     & surface          & regular         & syntactic               \\ \bottomrule
            corr\_adj\_var  & corrected adjective variation     & lexicalvariation & lexico-semantics & abstract        & lexical                 \\
            corr\_verb\_var & corrected verb variation          & lexicalvariation & lexico-semantics & abstract        & lexical                 \\ \bottomrule
            fkgl            & Flesch-Kincaid grade level        & readformula      & surface          & highly abstract & pragmatic               \\
            a\_kup\_pw      & Kuperman age of acquisition       & worddiff         & lexico-semantics & highly abstract & lexical                 \\ \bottomrule
        \end{tabular}%
    }
    \caption{Handcrafted Linguistic Features}\label{table-hlf}
\end{table*}
\footnotetext{as described in the figure in~\cite{lugea2023stylistics}, chapter 1, page 7}

@@

We finally look at the second variant of the second scenario of our experiment.
The hypothesis in this scenario variant was that the baseline would exhibit HLFs
that are much closer to the target corpus average if the input text is from that
corpus.
The expectation is to see fewer significant HLF results and that, as a rule of
thumb, HLF results are closer to the baseline in all cases.
The results shown in Table~\ref{table-prompt-2-ifd-cnn-dailymail} contradict our
expectations outright.
The most surprising behaviour is exhibited for HLFs which are derived into more
concrete generation instructions, such as the total number of words.
This behaviour is highlighted in figure~\ref{fig-p2-ifd-gemini-twords}.
Additionally, we notice regressive behaviour for some models compared to the
previously explored results in figure~\ref{fig-p2-ifd-claude3-nuadj}.
Finally, the abstractness of the text generation instructions for certain HLFs
leads to the generated text not being compliant as exemplified in
figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.

>>>>

\begin{table*}[ht!]
    \setlength\tabcolsep{6pt}
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{@{}llllll@{}}
            \toprule
            LFTK ID         & Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark\\ \toprule
            t\_word         & total words                       & wordsent         & surface          & very concrete   & lexical                 \\
            t\_sent         & total sentences                   & wordsent         & surface          & very concrete   & syntactic               \\ \bottomrule
            n\_uverb        & total number of unique verbs      & partofspeech     & syntax           & concrete        & syntactic               \\
            n\_uadj         & total number of unique adjectives & partofspeech     & syntax           & concrete        & syntactic               \\ \bottomrule
            simp\_ttr       & simple type token ratio           & typetokenratio   & lexico-semantics & regular         & pragmatic               \\
            a\_verb\_pw     & number of verbs per word          & partofspeech     & surface          & regular         & syntactic               \\ \bottomrule
            corr\_adj\_var  & corrected adjective variation     & lexicalvariation & lexico-semantics & abstract        & lexical                 \\
            corr\_verb\_var & corrected verb variation          & lexicalvariation & lexico-semantics & abstract        & lexical                 \\ \bottomrule
            fkgl            & Flesch-Kincaid grade level        & readformula      & surface          & highly abstract & pragmatic               \\
            a\_kup\_pw      & Kuperman age of acquisition       & worddiff         & lexico-semantics & highly abstract & lexical                 \\ \bottomrule
        \end{tabular}%
    }
    \caption{Handcrafted Linguistic Features}\label{table-hlf}
\end{table*}
\footnotetext{as described in the figure from~\cite{lugea2023stylistics}, chapter 1, page 7}

@@

We finally look at the second variant of the second scenario of our experiment.
The hypothesis in this scenario variant was that the baseline would exhibit HLFs
that are much closer to the target corpus average if the input text is from that
corpus.
The expectation is to see fewer significant HLF results and that, as a rule of
thumb, HLF results are closer to the baseline in all cases.
The results shown in Table~\ref{table-prompt-2-ifd-cnn-dailymail} contradict our
expectations outright.
The most surprising behaviour is exhibited for HLFs which are derived into more
concrete generation instructions, such as the total number of words.
This behaviour is highlighted in Figure~\ref{fig-p2-ifd-gemini-twords}.
Additionally, we notice regressive behaviour for some models compared to the
previously explored results in Figure~\ref{fig-p2-ifd-claude3-nuadj}.
Finally, the abstractness of the text generation instructions for certain HLFs
leads to the generated text not being compliant as exemplified in
Figure~\ref{fig-p2-ifd-llama3-a-kup-pw}.
